Testing Quasar Unification with Clumpy Wind Models

### 2015-11-12 paper.tex
minor edits, added 0946 back

Here, we continue our tests of  this paradigm by comparing ``average'' AGN spectra to 

@@

produce synthetic spectra in the rest-frame UV that have the characteristics of AGN, 
assuming total luminosities as high as 10$^{45}$ ergs s$^{-1}$. 

@@

allows for a more realistic AGN luminosities, while maintaining the ionization 

@@

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}). 
We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce strong BALs with $BI > ??$?
\item Does significant line emission emerge at low inclinations, with $EW\sim50\AA$ in \civ?
\item Do H recombination lines appear in the spectrum, $EW\sim50\AA$ in \la?
\item Do a certain range of angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
In this section, we present one of the most promising models and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

@@

% \begin{figure*} %fullpage
% \centering
% \includegraphics[width=1.0\textwidth]{figures/0946.png}
% \caption
% {
% Synthetic spectra at $72^\circ$ compared to an HST STIS spectrum
% of PG0946+301 (Arav et al. 2000). The spectrum is scaled to the flux at 1300\AA\
% for direct comparison. 
% }
% \label{fig:stis}
% \end{figure*} %fullpage

@@

The UV spectrum is characterised by strong BAL 
profiles at high inclinations ($> 70^\circ$). 
This highlights the first success of our model: 
clumping means the correct ionization state 
is maintained in the presence of strong X-rays, 
allowing large resonance line opacities. 
At the highest inclinations, the 
cooler, low ionization material at the base of the wind
starts to intersect the line of sight. This produces 
multiple absorption lines in species such as \mg,
\al\ and Fe~\textsc{ii}. The potential links to LoBALQSOs and 
FeLoBALQSOs are discussed in section 2.4.

@@

\subsubsection{The Angular Distribution of Line And Continuum Emission}
\label{angular}

@@

{\bf Note: One of the things I definitely wondered about is whether our "model" would produce the fraction of BALs seen in the sky}

@@

\begin{enumerate}
\item We have introduced a first-order treatment 
of clumping in our model, and found that it can now maintain
the required ionization state while agreeing well with the X-ray
properties of AGN/QSOs.
\smallskip
\item We have shown that the degree of ionization stratification
in the model is sufficient that LoBAL line profiles
are seen at a subset of viewing angles, and Fe~\textsc{ii}
absorption is seen at particularly high inclinations.
\smallskip
\item We find that clumping also causes a significant 
increase in the strength of the  emission
lines produced by the model. This is true both
of collisionally excited resonance lines (such as \civ, \nv)
and recombination lines (such as \la, \ha\ and the Balmer series).
\smallskip
\item The line EWs in our models increase with inclination.
BAL and non-BAL quasar composites have comparable EWs, so our model
fails to reproduce this behaviour.
This is due to a fundamental constraint discussed further in section 5. If the BLR
emits fairly isotropically then for a foreshortened, limb-darkened classical thin accretion disc
it is simply not possible to achieve line ratios at low inclinations that are comparable to
those at high inclinations. This is a robust conclusion which 
is independent of the assumed BLR geometry and size.
% \smallskip
% \item We have examined the effect of GR on our disc SED, using the disc atmosphere
% and GR ray-tracing code \agn. While including GR effects
% does cause the disc SED to become slightly more isotropic,
% the effect is not large enough to produce uniform line to continuum ratios
% with viewing angle. We briefly discuss other solutions.
\end{enumerate}
Our work confirms a number of expected outcomes from a geometric unification 
model, and suggests that a simple biconical geometry such as this can come close to 
explaining much of the  phenomenology of quasars. Nevertheless, our conclusions pose 
a clear challenge to the current disc wind unification picture.

>>>>

Here, we continue our tests of  this paradigm by comparing ``average'' quasar spectra to 

@@

produce synthetic spectra in the rest-frame UV that have the characteristics of quasars. 

@@

allows for a more realistic X-ray luminosities, while maintaining the ionization 

@@

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}). 
We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles?
\item Does significant line emission emerge at low inclinations, with $EW\sim40\AA$ in \civ?
\item Do H recombination lines appear in the spectrum, $EW\sim50\AA$ in \la?
\item Do a small subset of BAL angles produce LoBAL features?
\item Does the model compare favourably to quasar composite spectra?
\end{itemize}
In this section, we present one of the most promising models and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/0946_estfix.png}
\caption
{
Synthetic spectra at $72^\circ$ compared to an HST STIS spectrum
of PG0946+301 (Arav et al. 2000). The spectrum is scaled to the flux at 1300\AA\
for direct comparison. 
}
\label{fig:stis}
\end{figure*} %fullpage

@@

The UV spectrum is characterised by strong BAL 
profiles at high inclinations ($> 70^\circ$). 
This highlights the first success of our model: 
clumping means the correct ionization state 
is maintained in the presence of strong X-rays, 
allowing large resonance line opacities. 
At the highest inclinations, the 
cooler, low ionization material at the base of the wind
starts to intersect the line of sight. This produces 
multiple absorption lines in species such as \mg,
\al\ and Fe~\textsc{ii}. The potential links to LoBALQSOs and 
FeLoBALQSOs are discussed in section 2.4.
Clearly, the absorption lines in the 
synthetic spectra at BALQSO-like angles do not match the composites; this
is {\em not} due to a failure of our model, but rather because a geometric mean
BALQSO composite spectrum tends to wash out the broad absorption features due to the wide
range of absorption characteristics. To demonstrate this, we show a comparison 
to a {\sl Hubble Space Telescope} STIS spectrum of the high BALnicity BALQSO PG0946+301 
(Arav et al. 2000) in figure \ref{fig:stis}.

@@

\subsubsection{The Angular Distribution of Line And Continuum Emission}
\label{{sec:angular}}

@@

@@

\begin{enumerate}
\item We have introduced a first-order treatment 
of clumping in our model, and found that it can now maintain
the required ionization state while agreeing well with the X-ray
properties of AGN/QSOs.
\smallskip
\item We have shown that the degree of ionization stratification
in the model is sufficient that LoBAL line profiles
are seen at a subset of viewing angles, and Fe~\textsc{ii}
absorption is seen at particularly high inclinations.
\smallskip
\item We find that clumping also causes a significant 
increase in the strength of the  emission
lines produced by the model. This is true both
of collisionally excited resonance lines (such as \civ, \nv)
and recombination lines (such as \la, \ha\ and the Balmer series).
\smallskip
\item The line EWs in our models increase with inclination.
BAL and non-BAL quasar composites have comparable EWs, so our model
fails to reproduce this behaviour.
This is due to a fundamental constraint discussed further in section~\ref{sec:angular}. If the BLR
emits fairly isotropically then for a foreshortened, limb-darkened classical thin accretion disc
it is simply not possible to achieve line ratios at low inclinations that are comparable to
those at high inclinations. This is a robust conclusion which 
is independent of the assumed BLR geometry and size.
% \smallskip
% \item We have examined the effect of GR on our disc SED, using the disc atmosphere
% and GR ray-tracing code \agn. While including GR effects
% does cause the disc SED to become slightly more isotropic,
% the effect is not large enough to produce uniform line to continuum ratios
% with viewing angle. We briefly discuss other solutions.
\end{enumerate}
Our work confirms a number of expected outcomes from a geometric unification 
model, and suggests that a simple biconical geometry such as this can come close to 
explaining much of the  phenomenology of quasars. Nevertheless, our conclusions pose 
a clear challenge to the current disc wind unification picture.
